---
title: "Seurat analysis report of BD Rhapsody TAS-Seq WTA data"
date: "`r Sys.Date()`"
author: "SS"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
    thumbnails: false
    lightbox: false
    gallery: false
    mathjax: false
    toc_depth: 2  #type-of-column header specification 
    md_extensions: -ascii_identifiers
params:
  cDNA_matrix: cDNA_matrix
  Tag_matrix: Tag_matrix
  nTags: nTags
  threads: threads
  DBEC_output: DBEC_output
  species: species
  min.ave: min.ave
  min.diff: min.diff
  SeqGeq_output: SeqGeq_output

---

```{r setup, echo=FALSE, cache=FALSE, message=FALSE, warning=FALSE}
setwd("../")
suppressWarnings(suppressMessages(source("./Rscripts/library_source_Seurat.R")))
suppressMessages(library(knitr))
suppressMessages(library(rmdformats))
suppressMessages(library(stringr))
suppressMessages(library(DT))
suppressMessages(library(cluster))
suppressMessages(library(parallelDist))
suppressMessages(library(fastSave))


quiet <- function(x) { 
  sink(tempfile()) 
  on.exit(sink()) 
  invisible(force(x)) 
} 

tableread_fast_hashtag = function(x, sep="\t", header=TRUE){
  tmp = data.table::fread(x, header=header, sep=sep, quote="")
  tmp = as.data.frame(tmp)
  return(tmp)
}

custom_colors <- list()
colors_dutch <- c(
  '#FFC312','#C4E538','#12CBC4','#FDA7DF','#ED4C67',
  '#F79F1F','#A3CB38','#1289A7','#D980FA','#B53471',
  '#EE5A24','#009432','#0652DD','#9980FA','#833471',
  '#EA2027','#006266','#1B1464','#5758BB','#6F1E51'
)
colors_spanish <- c(
  '#40407a','#706fd3','#f7f1e3','#34ace0','#33d9b2',
  '#2c2c54','#474787','#aaa69d','#227093','#218c74',
  '#ff5252','#ff793f','#d1ccc0','#ffb142','#ffda79',
  '#b33939','#cd6133','#84817a','#cc8e35','#ccae62'
)
colors_custom = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c',
                      '#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#b15928',
                      '#49beaa', '#611c35', '#2708a0')
custom_colors$discrete <- unique(c(colors_dutch, colors_spanish, colors_custom))


## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               fig.align='center')
opts_knit$set(width=50)

cDNA_matrix = params$cDNA_matrix
Tag_matrix = params$Tag_matrix
nTags = as.numeric(params$nTags)
threads = as.numeric(params$threads)
DBEC_output = as.logical(params$DBEC_output)
species = params$species
min.ave = as.numeric(params$min.ave)
min.diff = as.numeric(params$min.diff)
SeqGeq_output = as.logical(params$SeqGeq_output)
sample_name = gsub("matrix_inflection_", "", cDNA_matrix)
sample_name = gsub(".txt.gz", "", sample_name)

poisson_notEmpty = 1-dpois(0, lambda=20000/200000)
poisson_singlet = dpois(1, lambda=20000/200000)
pct.doublet = 1-poisson_singlet/poisson_notEmpty
pct.doublet.detectable = pct.doublet*(choose(nTags+2-1, 2)-nTags)/choose(nTags+2-1, 2)

#create output folders
#make output directories
dir.name = sprintf("./result/%s_results/", sample_name)
dir.name.1 = paste(dir.name, "Seurat", sep="")
if(!dir.exists(dir.name.1)){
dir.create(dir.name.1)
}

dir.name.plot = paste("./result/", sample_name, "_results/plots/", sep="")
dir.name.matrix = paste("./result/", sample_name, "_results/matrix/", sep="")
dir.name.plot1 = paste("./result/", sample_name, "_results/Seurat/Seurat_plots/", sep="")

if(!dir.exists(dir.name.plot)){
dir.create(dir.name.plot)
}

if(!dir.exists(dir.name.plot1)){
dir.create(dir.name.plot1)
}

if(species== "macaca"){
  species = "hsa"
}


```

## Brief summary of the demultiplexing procedures

For assignment of each tags to each cell barcodes, read counts of each tag in each valid cell barcode, which defined by the cDNA matrix, were extracted from tag/cell barcode expression matrix. Unassigned  cell barcodes were labeled as "not-detected" cells. Then, sum of the total read counts of each tags were normalized to the minimum sum count of each tags, and log2 fold-change between first most tag counts and second most tag counts within each cell barcode. Each cell barcode were ranked by the fold-change ascending order, and top `r paste(round(pct.doublet.detectable*100, digits=2), "% cells", sep="")` were identified as doublets, which were theoretically detectable doublets calculated by the Poisson's distribution based on the number of loaded cells, total Rhapsody well number, and number of tags used.

Finally, remained cell barcodes were assigned to the first most counted tags. Each tag expression in each cell barcode were log2(x+1)-transformed, z-scaled by each cell barcode, and visualized by using pheatmap package in R 3.6.3.  


```{r analysis_demultiplex, echo=FALSE, cache=FALSE, message=FALSE}

#create output folders
setwd("../") #must set working directory. default is .Rmd-existing path.


#load gene expression matrix
tablelist = lapply(paste("./result/", sample_name, 
                         "_results/matrix/", cDNA_matrix, sep=""), 
                   tableread_fast_sparse)
names(tablelist)=sample_name

#load hashtag-expression matrix
hashtag_fnames1 = gsub("Hashtag_top1M_", "", Tag_matrix)
hashtag_fnames1 = gsub(".txt.gz", "", hashtag_fnames1)
hashtag_data=lapply(paste("./result/", hashtag_fnames1, "_results/matrix/", Tag_matrix, sep=""), tableread_fast_hashtag)
hashtag_data[[1]][,1] = paste(names(tablelist)[1], hashtag_data[[1]][,1], sep='_')
quiet(gc())


##create Seurat object (Seurat v2 workflow) and annotate by hashtag
colnames(tablelist[[1]]) = paste(names(tablelist)[1], colnames(tablelist[[1]]), sep='_')
seu = lapply(tablelist, CreateSeuratObject, min.cells = 5, min.genes = 200)

dir.name.plot = paste("./result/", sample_name, "_results/plots/", sep="")
dir.name.matrix = paste("./result/", sample_name, "_results/matrix/", sep="")
dir.name.plot1 = paste("./result/", sample_name, "_results/Seurat/Seurat_plots/", sep="")

#Add hashtag annotation metadata
seu = Demultiplex_DNAtags(hashtag_data[[1]], seu[[1]],
                          scale.factor = max(colSums(hashtag_data[[1]][,2:ncol(hashtag_data[[1]])])),
                                 nTags=nTags, nCells_in_cartridge=nrow(seu[[1]]@meta.data), 
                                 sample.name=sample_name[[1]],
                                 dir.name=dir.name.plot)
hoge = paste(dir.name.plot, sample_name, "_demultiplex.txt", sep="")
quiet(file.remove(hoge))

#add hashtag ID to cell_barcodes and export expression matrix 
raw.data = seu[[1]]@raw.data
tmp = paste(seu[[1]]@meta.data$TagIDs, rownames(seu[[1]]@meta.data), sep="_")
colnames(raw.data)=tmp
raw.data=as.matrix(raw.data)
file.name.raw = sprintf("matrix_inflection_demulti_%s.txt.gz", sample_name)
file.name.raw = paste(dir.name.matrix, file.name.raw, sep="")
fwrite(as.data.frame(raw.data), file.name.raw, row.names=T, col.names=T, sep="\t", 
       eol="\n", quote=F, nThread=threads, compress="gzip")

tmp = table(seu[[1]]@meta.data$TagIDs)
tmp = as.data.frame(tmp)
rownames(tmp)=tmp[,1]
logs_Ncells = as.data.frame(t(tmp))
logs_Ncells = logs_Ncells[2,,drop=F]
rownames(logs_Ncells)="cell number"

#add normalized hashtag exporession data to gene_expression matrix for SeqGeq software
if(SeqGeq_output){
hashtag_expression_data = seu[[1]]@meta.data[,5:(nTags+4)]
rownames(hashtag_expression_data)=colnames(raw.data)
raw.data1 = rbind(raw.data, t(2^hashtag_expression_data))
file.name = sprintf("matrix_demulti_SeqGeq_%s.txt.gz", sample_name) 
file.name = paste(dir.name.matrix, file.name, sep="")
fwrite(as.data.frame(raw.data1), file.name, row.names=T, col.names=T, sep="\t", 
       eol="\n", quote=F, nThread=threads, compress="gzip")
}

```

---

Figure 2: Heatmap representation of tag expression and assignment in each cell barcodes

```{r output_demultiplex_plots, echo=FALSE, cache=FALSE, fig.width=8, fig.height=4, dpi=300, out.width = 800, out.height = 400, message=FALSE, fig.align='center'}
plot_grid(seu[[2]])

```

```{r output_table1, echo=FALSE, cache=FALSE,dpi=300, out.width = 1000, out.height = 500, fig.align='center'}
DT::datatable(logs_Ncells, filter="none", width='100%',
          caption = 'Table 1: Number of cells assigned to each Tags', 
          class = 'cell-border stripe', 
          options = list(pageLength=1, searching=FALSE, lengthChange=FALSE, autoWidth = TRUE, scrollX=F)) %>%
   DT::formatStyle(columns = c(1:ncol(logs_Ncells)), fontSize = 8)
```

File locations  

Demultiplexed plot was stored at `r paste(dir.name.plot, sprintf("%s_demultiplex.png", sample_name), sep="")`.
Demultiplexed expression matrix was stored at `r paste(dir.name.matrix, sprintf("matrix_inflection_demulti_%s.txt.gz", sample_name), sep="")`.
`r if(SeqGeq_output){paste("Demultiplexed expression matrix for SeqGeq analysis was stored at ", dir.name.matrix, sprintf("matrix_demulti_SeqGeq_%s.txt.gz", sample_name), sep="")}`

---

`r if(DBEC_output){"Perform DBEC correction (mRNA leak-derived backgroud subtraction of gene expression)"}`
`r if(DBEC_output){paste("For DBEC correction, genes of which log2(x+1)-transformed maximum expression over 8 were selected, and biexponential transformation was performed by using FlowTrans package in R 3.6.3. Next, normal mixture components (model E) was detected by using mclust package in R 3.6.3, average expression of each components was calculated, and genes of which the maximum average expression over ", min.ave, " were selected for background subtration. Then, if the components of which difference of the maximum average expression against maximal average expression under ", min.diff, " were detected, the expression value of the components set as 0 for background subtracion.", sep="")}`


```{r analysis_DBEC, echo=FALSE, cache=FALSE, eval = DBEC_output}

setwd("../") #must set working directory. default is .Rmd-existing path.
message("Performing DBEC-correction... This may take some time.")

 DBEC_filter = background_subtraction_Biex(tablelist, min.event=100, minimum.max.expr=8,
                                          species=species, min.ave=min.ave, AutoThreshold = FALSE,
                                          min.diff=min.diff, modelnames="E",
                                          uncert.thre=1, nthreads=threads, sample.name=sample_name,
                                          dir.name=dir.name.plot)
 names(DBEC_filter) = sample_name
 DBEC_res = apply_DBEC_filter(tablelist, DBEC_filter=DBEC_filter, nthreads=threads, sample.name = sample_name)
 names(DBEC_res) = sample_name

 ##create Seurat object (Seurat v2 workflow) and annotate by hashtag
 seu = lapply(DBEC_res, CreateSeuratObject, min.cells = 5, min.genes = 200)

 #Add hashtag annotation metadata
 seu = Demultiplex_DNAtags(hashtag_data[[1]], seu[[1]],
                           scale.factor = max(colSums(hashtag_data[[1]][,2:ncol(hashtag_data[[1]])])),
                               nTags=nTags, nCells_in_cartridge=nrow(seu[[1]]@meta.data), 
                               sample.name=sample_name[[1]],
                               dir.name=dir.name.plot)
 hoge = paste(dir.name.plot, sample_name, "_demultiplex.txt", sep="")
 quiet(file.remove(hoge))

 #add hashtag ID to cell_barcodes and export expression matrix 
raw.data = seu[[1]]@raw.data
tmp = paste(seu[[1]]@meta.data$TagIDs, rownames(seu[[1]]@meta.data), sep="_")
colnames(raw.data)=tmp
colnames(raw.data)=gsub("not_detected", "not-detected", colnames(raw.data))
raw.data=as.matrix(raw.data)
file.name.raw = sprintf("matrix_inflection_demulti_DBEC_%s.txt.gz", sample_name)
file.name.raw = paste(dir.name.matrix, file.name.raw, sep="")
fwrite(as.data.frame(raw.data), file.name.raw, row.names=T, col.names=T, sep="\t", 
       eol="\n", quote=F, nThread=threads, compress="gzip")

 tmp = table(seu[[1]]@meta.data$TagIDs)
 tmp = as.data.frame(tmp)
 rownames(tmp)=tmp[,1]
 logs_Ncells = as.data.frame(t(tmp))
 logs_Ncells = logs_Ncells[2,,drop=F]
 rownames(logs_Ncells)="cell number"

 
 #add normalized hashtag exporession data to gene_expression matrix for SeqGeq software
 if(SeqGeq_output){
 hashtag_expression_data = seu[[1]]@meta.data[,5:(nTags+4)]
 rownames(hashtag_expression_data)=colnames(raw.data)
 raw.data1 = rbind(raw.data, t(2^hashtag_expression_data))
 file.name = sprintf("matrix_inflection_demulti_DBEC_SeqGeq_%s.txt.gz", sample_name)
 file.name = paste(dir.name.matrix, file.name, sep="") 
 fwrite(as.data.frame(raw.data1), file.name, row.names=T, col.names=T, sep="\t", 
       eol="\n", quote=F, nThread=threads, compress="gzip")
 }
 

```

`r if(DBEC_output){paste("File locations", sep="")}`  
`r if(DBEC_output){paste("DBEC threshold plots were stored at ", dir.name.plot, sprintf("DBEC_thresholds_%s", sample_name), sep="")}`
`r if(DBEC_output){paste("Demultiplexed/DBEC-corrected expression matrix was stored at ", dir.name.matrix, sprintf("matrix_inflection_demulti_DBEC_%s.txt.gz", sample_name), sep="")}`  
`r if(DBEC_output && SeqGeq_output){paste("Demultiplexed/DBEC-corrected expression matrix for SeqGeq analysis was stored at ", dir.name.matrix, sprintf("matrix_inflection_demulti_DBEC_SeqGeq_%s.txt.gz", sample_name), sep="")}`  
`r if(DBEC_output){paste("Figure 2: Heatmap representation of tag expression and assignment in each cell barcodes (after DBEC correction)", sep="")}`


```{r output_demultiplex_plots2, echo=FALSE, cache=FALSE, eval = DBEC_output, fig.width=8, fig.height=4, dpi=300, out.width = 800, out.height = 400, message=FALSE, fig.align='center'}
plot_grid(seu[[2]])

```


```{r output_table2, echo=FALSE, cache=FALSE, eval = DBEC_output, dpi=300, out.width = 1000, out.height = 500, fig.align='center'}
DT::datatable(logs_Ncells, filter="none", width='100%',
          caption = 'Table 1: Number of cells assigned to each Tags (DBEC-corrected matrix)', 
          class = 'cell-border stripe', 
          options = list(pageLength=1, searching=FALSE, lengthChange=FALSE, autoWidth = TRUE, scrollX=F)) %>%
   DT::formatStyle(columns = c(1:ncol(logs_Ncells)), fontSize = 8)
```

```{r Seurat_v2_step1, echo=FALSE, cache=FALSE}

quiet(gc())
setwd("../") #must set working directory. default is .Rmd-existing path.

 #re-generate Seurat object for demultiplexed sample
 raw.data = seu[[1]]@raw.data
 tmp = paste(seu[[1]]@meta.data$TagIDs, rownames(seu[[1]]@meta.data), sep="_")
 colnames(raw.data)=as.character(tmp)
 colnames(raw.data)=gsub("not_detected", "not-detected", colnames(raw.data))
 seu=NULL
 seu = CreateSeuratObject(raw.data, min.cells=5, min.genes=200)
 colnames(seu@meta.data)[2]="nReads"
 file.name.Seurat=paste(dir.name, "/Seurat/", sample_name, "_Seurat.rda", sep='')
 save.pigz(seu, file=file.name.Seurat, n.cores=threads*2)

 #make cell name-cell barcode table for RNA velocity analysis
fuga = unique(seu@meta.data$orig.ident)
fuga1 = paste(fuga, "_", sample_name, "_", sep="")

hoge1 = rownames(seu@meta.data)
for (i in c(1:length(fuga))){
 hoge1 = gsub(fuga1[i], "", hoge1)
}

res = data.frame(Barcode=hoge1, Nreads_CellBC=rownames(seu@meta.data))
hoge2 = paste(dir.name.matrix, sample_name, ".csv", sep="")
write.table(res, hoge2, row.names=F, col.names=T, sep=",", quote=F, append=F)

```

File location   

Entire Seurat object file was stored at `r paste(file.name.Seurat, sep="")`.

